### 2014-04-01 paper/relatedwork.tex
shortening

\section{Related Work}

\tovi{make sure to have mini intros and conclusions for each subsection.  it's ok to discuss several papers at once if they are similar.  each paper needs a sentence to say what distinguishes us from it}

Our contributions relate to other efforts to add interactivity to digitally fabricated objects; and to techniques for routing and model modification.  \bjoern{The related work section is taking shape but still has a bit of a "missing the forest for the trees" feeling.)}

\bjoern{Tactically, to communicate who would be is a qualified reviewer for this paper in the UIST community, we should make sure to have appropriate references to Baudisch, Poupyrev, Harrison in here}

\valkyrie{can you offer suggestions?  I can imagine the latex buttons paper, touche, acoustic barcodes, capacitive fingerprinting \cite{Harrison-capfingerprint}, printed optics for harrison/poupyrev, and for baudisch maybe.... TDR \cite{Wimmer-TDR}? fiberio \cite{Holz-fiberio}? fabrickation? I'm not sure on this.}

@@

\bjoern{May be able to condense this entire subsection into ~10 lines}
Other research has attempted to bring traditional electronics to 3D prints, either on the surface or via interior channels.  Sells, et al., in \cite{Sells-reprap} create channels on the surface of a 3D printed object which are filled with a special computer-controlled syringe of liquid solder to create connections.  Their focus is on creating self-replicating 3D printers, and their electronics are confined to traditional 2D routed circuitboards.  Sarik, et al., \cite{Sarik-tracebrush} \valkyrie{I never found a real paper of this.  The Microsoft website has a "print" of it with TEI boilerplate, but I couldn't find it in the digital library.  What's up with that?} and the Optomec company \cite{optomec} both created techniques for taking an already-printed object and spraying conductive material onto its surface.  In contrast, our technique can lead to more durable circuitry, since it is routed through the core of the object rather than on its surface.  Navarette, et al., in \cite{Navarrette-gps} created a functioning GPS whose circuit routing extended into true 3D (as compared to the 2.5D boards created by stacking traditional 2D circuitboards).  However, they created this one-off routing by hand and do not offer any general algorithms or techniques.  In the area of robotics, softer-than-skin interactive surfaces have been created using tubed interfaces fabricated from silicone molds.  \cite{Park-microchannels} and \cite{Majidi-curvature} use microchannels filled with liquid metal for pressure or curvature sensing: the metal in the channels changes resistance as the channels are stretched.  These techniques have not been used with fabrication, nor has a design tool to create such sensors been previously described.

@@

Circuitboard routing algorithms, such as Lee's maze router algorithm \bjoern{ref?}, work only on a 2D surface. Multi-layer boards are typically 2.5D, in that they comprise a stack of 2D circuitboards connected by vias \valkyrie{citation needed}.  Instead we offer true 3d point to point routing though the interior of solid models.

@@

There is prior work on modeling the interior of objects (as opposed to their exterior).  In Make it Stand, material is removed from the interior of objects with the goal of changing their center of gravity \cite{Prevost-makeitstand}.  The creation of such hollow chambers is a part of our design space, however our contribution is orthogonal as we do not focus on re-balancing objects using these chambers.  Bickel, et al., explored enforcing deformation behavior with material blending \cite{Bickel-deformation}, while B\"{a}cher, et al., built an algorithm for automatically placing 3D fabricatable joints into meshes \cite{Bacher-articulated}.  Our focus is not on articulated or deformable 3D prints, although we do perform interior modeling.  Takayama, et al., built a system described in \cite{Takayama-fruit} which allowed users to create internal color maps of the interior of objects, allowing them to see ``slices'' of their ``oranges.''  This differs from our contribution, as our pipes are designed for fabrication, and focus on function rather than appearance.  Sauron automatically modified the interior of existing mechanical objects to make them amenable to vision-based sensing \cite{Savage-sauron}, and Infrastructs created encoded interior cavities in 3D printed objects designed to be sensed with a terahertz imaging tool.  Our work differs in that we want to allow users to design pipes and integrate interactive objects with them. \bjoern{Can condense this - they all make orthogonal contributions, so don't need to repeat this statement of differentiation n times. Just do it once at the end.}

>>>>

\section{Related Work}
Our contributions relate to other efforts to add interactivity to digitally fabricated objects; and to techniques for routing and model modification.

@@

Other research has brought traditional electronics to 3D prints.  Sells, et al., \cite{Sells-reprap}, describe using a heated syringe to deposit metal in printed 2D circuitboard channels on objects' surfaces.  Sarik, et al., \cite{Sarik-tracebrush} \valkyrie{I never found a real paper of this.  The Microsoft website has a "print" of it with TEI boilerplate, but I couldn't find it in the digital library.  What's up with that?} and the Optomec company \cite{optomec} describe \emph{spraying} conductive material onto objects to create circuitry.  Injection of liquid metals for sensing in softer-than-skin electronics was explored by Park, et al., \cite{Park-microchannels} and Majidi, et al., \cite{Majidi-curvature}.  Navarette, et al., created a functioning GPS routed in true 3D (not multilayer boards' 2.5D), hand-built using a one-off routing.  Our work differs from these because we offer \emph{algorithms and techniques} for \emph{internal}, \emph{true 3D} routing.

@@

Circuitboard routing algorithms, such as Lee's maze router algorithm \cite{Lee-maze}, work only on a 2D surface. Multi-layer boards are typically 2.5D, in that they comprise a stack of 2D circuitboards connected by vias.  Instead we offer true 3d point to point routing though the interior of solid models.

@@

There is prior work on modeling the interior of objects (as opposed to their exterior).  In Make it Stand, material is removed from the interior of objects with the goal of changing their center of gravity \cite{Prevost-makeitstand}.  Bickel, et al., explored enforcing deformation behavior with material blending \cite{Bickel-deformation}, while B\"{a}cher, et al., built an algorithm for automatically placing 3D fabricatable joints into meshes \cite{Bacher-articulated}.  Takayama, et al., built a system described in \cite{Takayama-fruit} which allowed users to create internal color maps of the interior of objects, allowing them to see digital ``slices'' of their ``oranges.''  Sauron automatically modified the interior of existing mechanical objects to make them amenable to vision-based sensing \cite{Savage-sauron}, and Infrastructs created encoded interior cavities in 3D printed objects designed to be sensed with a terahertz imaging tool.  All these contributions are orthogonal to ours, as our focus is on modeling pipes and hollow chambers for fabricating interactive objects.
